On the dynamic assessment of the Local-Maximum Entropy Material Point Method through an Explicit Predictor-Corrector Scheme

### 2020-09-03 LME-MPM.tex
Nuevo caso (choque de una placa)

\MODIFIED{First, the velocity field and the virtual displacements fields are discretised in a finite set of nodes $\textbf{X} = \{ \vec{x}_I, I \in \mathcal{B} \} \subset \mathbb{R}^n$, where $\mathcal{B} = 1, \ldots, N_I$. The continuum field can be reconstructed with the help of nodal values and an appropriate interpolation function $N_I(\vec{x})$. Also spatial derivatives of those quantities, such as gradients and divergences, are computed through the support of the background set of nodes as.}
\begin{align}
    \label{eq:variable_reconstruction}
    \varphi(\vec{x}) &= \sum_{I\ \in\ \mathcal{B}} N_I(\vec{x})\ \varphi_I \\
    \label{eq:grad_variable_reconstruction}
    \Grad{\varphi}(\vec{x}) &= \sum_{I\ \in\ \mathcal{B}} \Grad{N_I(\vec{x})}\ \varphi_I
\end{align}
\MODIFIED{Secondly, the continuum \gls{domain} is discretised with a finite set of material points (also denominated particles the manuscript) $\hat{\Omega} = \{ \vec{x}_p, p \in \mathcal{C} \} \subset \Omega$, where $\mathcal{C} = 1, \ldots, N_p$. Any particle field such as position, velocity, mass, volume and stress denoted by $\vec{x}_p$, $\vec{v}_p$, $m_p$, $V_p$ and $\tens{\sigma}_p$ respectively are assigned to each material point. Furthermore, any particle field $\varphi_p$ can be reconstructed within the nodes in the neighbourhood of each particle, $\mathcal{B}_p \subset \mathcal{B}$, and evaluating the interpolation function in the position of each particle as}
\begin{align}
    \label{eq:particle_variable_reconstruction}
\varphi_p &= \sum_{I\ \in\ \mathcal{B}_p} N_I(\vec{x}_p)\ \varphi_I \\
\Grad{\varphi_p} &= \sum_{I\ \in\ \mathcal{B}_p} \Grad{N_I(\vec{x}_p)}\ \varphi_I
\end{align}
\MODIFIED{Regarding numerical integration, the material derivatives are evaluated using the Riemann integral definition \cite{Riemann_1854} applied to a finite set of points which associated volumes $V_p$ are interpreted as quadrature weights.}
\begin{equation}
    \label{eq:particle_riemann_integral}
\Integral{3}{\varphi} = \sum_{p\ \in\ \mathcal{C}} \varphi_p\ V_p 
\end{equation}

@@

\MODIFIED{Let us illustrate the procedure described above by fully developing the term of the acceleration forces in \eqref{eq:BalanceMomentum_wf}. Employing the definition \eqref{eq:variable_reconstruction} the velocity and virtual displacement fields can be approximated within its nodal values. Hence, the acceleration forces is reduced to}

@@

\MODIFIED{Where $\tens{m}_{IJ}$, is the nodal mass matrix, and $N_{Ip}, N_{Jp}$ are the interpolation functions for nodes $I,J$ evaluated in the position of each particle $p$. In order to improve the computational efficiency and stability, the nodal mass matrix can be substituted by the lumped mass matrix $\tens{m}_{IJ}^{lumped}$. The remaining terms of \eqref{eq:BalanceMomentum_wf} are obtained within a similar procedure and can be found in literature, see for instance \cite{Zhang_book_2016}.

@@

This section is devoted to test the ability of both predictor-corrector
time integration scheme and the Local \textit{Max-Ent} approximants to
overcome spurious oscillations due to the grid crossing and high
frequency loads under the context of \acrshort{mpm}. \MODIFIED{Three different test have
been adopted for this purpose: the benchmark proposed by Dyka \& Ingel
(1995)\cite{Dyka1995}, the test proposed in the PhD thesis of
Andersen (2009)\cite{thesis_Andersen_2009} and the classical example of two colliding discs provided by Sulsky {\it et al.} (1994) \cite{Sulsky1994}. The first example is devoted to test the accuracy of the \acrfull{npc} scheme is compared to the standard \acrfull{fe} in order to assess the performance of the time integration schemes. In the second example,  \acrshort{lme} solutions
are compared with those provided by \acrfull{ugimp} and Q4 shape
functions. Finally, in the third example the present approach is tested under a scenario where shocks and grid crossing occurs.} To avoid some mesh-dependent issues, in all
calculations a regular background mesh was set. All simulations were
performed with in-house software.

@@

\subsection{Impact of two elastic discs}
\label{sec:Impact-two-elastic-discs}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\sidecaption
 \centering
 \resizebox{0.7\hsize}{!}{
   \input{Figures/Sulsky_discs.tex}}
 \caption{Geometrical description of the two colliding discs problem.}
 \label{fig:Sulsky_discs}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

\MODIFIED{First, the velocity field and the virtual displacements fields are discretised in a finite set of nodes $\textbf{X} = \{ \vec{x}_I, I \in \mathcal{B} \} \subset \mathbb{R}^n$, where $\mathcal{B} = 1, \ldots, N_I$. The continuum field can be reconstructed with the help of nodal values and an appropriate interpolation function $N_I(\vec{x})$. Also spatial derivatives of those quantities, such as gradients and divergences, are computed through the support of the background set of nodes as}
\begin{align}
    \label{eq:variable_reconstruction}
    \varphi(\vec{x}) &= \sum_{I\ \in\ \mathcal{B}} N_I(\vec{x})\ \varphi_I \\
    \label{eq:grad_variable_reconstruction}
    \Grad{\varphi}(\vec{x}) &= \sum_{I\ \in\ \mathcal{B}} \Grad{N_I(\vec{x})}\ \varphi_I
\end{align}
\MODIFIED{Secondly, the continuum \gls{domain} is discretised with a finite set of material points (also denominated particles the manuscript) $\hat{\Omega} = \{ \vec{x}_p, p \in \mathcal{C} \} \subset \Omega$, where $\mathcal{C} = 1, \ldots, N_p$. Any particle field such as position, velocity, mass, volume and stress denoted by $\vec{x}_p$, $\vec{v}_p$, $m_p$, $V_p$ and $\tens{\sigma}_p$ respectively are assigned to each material point. Furthermore, any particle field $\varphi_p$ can be reconstructed within the nodes in the neighbourhood of each particle, $\mathcal{B}_p \subset \mathcal{B}$, and evaluating the interpolation function in the position of each particle as}
\begin{align}
    \label{eq:particle_variable_reconstruction}
\varphi_p &= \sum_{I\ \in\ \mathcal{B}_p} N_I(\vec{x}_p)\ \varphi_I \\
\Grad{\varphi_p} &= \sum_{I\ \in\ \mathcal{B}_p} \Grad{N_I(\vec{x}_p)}\ \varphi_I
\end{align}
\MODIFIED{Regarding numerical integration, the material derivatives are evaluated using the Riemann integral definition \cite{Riemann_1854} applied to a finite set of points which associated volumes $V_p$ are interpreted as quadrature weights.}
\begin{equation}
    \label{eq:particle_riemann_integral}
\Integral{3}{\varphi} = \sum_{p\ \in\ \mathcal{C}} \varphi_p\ V_p 
\end{equation}

@@

\MODIFIED{Let us illustrate the procedure described above by fully developing the term of the acceleration forces in \eqref{eq:BalanceMomentum_wf}. Employing the definition \eqref{eq:variable_reconstruction}, the velocity and virtual displacement fields can be approximated within its nodal values. Hence, the acceleration forces is reduced to}

@@

\MODIFIED{$\tens{m}_{IJ}$, is the nodal mass matrix, and $N_{Ip}, N_{Jp}$ are the interpolation functions for nodes $I,J$ evaluated in the position of each particle $p$. In order to improve the computational efficiency and stability, the nodal mass matrix can be substituted by the lumped mass matrix $\tens{m}_{IJ}^{lumped}$. The remaining terms of \eqref{eq:BalanceMomentum_wf} are obtained within a similar procedure and can be found in literature, see for instance \cite{Zhang_book_2016}.

@@

This section is devoted to test the ability of both predictor-corrector
time integration scheme and the Local \textit{Max-Ent} approximants to
overcome spurious oscillations due to the grid crossing and high
frequency loads under the context of \acrshort{mpm}. \MODIFIED{Three different tests have
been adopted for this purpose: the benchmark proposed by Dyka \& Ingel
(1995)\cite{Dyka1995}, the test proposed in the PhD thesis of
Andersen (2009)\cite{thesis_Andersen_2009} and the evolution of velocity waves in an elastic square. The first example is devoted to test the accuracy of the \acrfull{npc} scheme is compared to the standard \acrfull{fe} in order to assess the performance of the time integration schemes. In the second example,  \acrshort{lme} solutions
are compared with those provided by \acrfull{ugimp} and Q4 shape
functions. Finally, in the third example the present approach is tested under a scenario where shocks and grid crossing occurs.} To avoid some mesh-dependent issues, in all
calculations a regular background mesh was set. All simulations were
performed with in-house software.

@@

\MODIFIED{\subsection{Velocity waves in an elastic domain}}
\label{sec:Velocity-waves-elastic-domain}

\MODIFIED{The following test is best suited to validate the ability of the discussed
time integration scheme to reproduce 2D wave propagation problem.  It consists of the side impact of a moving elastic square. The square was 50 x 50 mm and modeled with the following elastic parameters :
\begin{itemize} 
\item  Initial density : $20$
\item  Poisson ratio : $0.3$
\item  Elastic modulus : $10^5$
\end{itemize}
}

\MODIFIED{The background set of nodes consists in a cartesian grid with a uniform gap between nodes of 1 mm, and square was discretised with four material points by cell. The courant number is fixed to 0.1 during the whole simulation and $\text{LME}_{4.0}$ shape functions where used. The square initially move horizontally through the grid at $\text{V}_0 = 1.0$ until it hits an obstacle modelled as a set of nodes with  zero velocity prescribed. The collision produce stress waves and cause the square to bounce back.}

\MODIFIED{Snapshots of velocity field magnitude within the square are given in\ref{fig:Magnitude_velocity_impact_square}. At early times, \acrshort{npc} and \acrshort{fe} are similar. At later times, the \acrshort{fe} introduces significant numerical noise while \acrshort{npc} is almost clean. }
%%%%%%%%%%%%%%%%
\begin{figure}
%\sidecaption
  \centering
  \resizebox{\hsize}{!}{
\includegraphics[scale=1]{Figures/Velocity_FE_vs_NPC_square.png} 
  }
  \caption{Snapshots of velocity magnitudes in an elastic square impacted at a point on the right edge. The top row is for \acrshort{npc} updates and the bottom row is for \acrshort{fe} updates.}
  \label{fig:Magnitude_velocity_impact_square}
\end{figure}
%%%%%%%%%%%%%%%% 
\MODIFIED{Energy plots in figure \ref{fig:Energy_impact_square} shows that after collision the total amount of energy is preserved. In the bounce back stress wave are still travelling thorough the elastic domain introducing numerical noise due to the inaccuracies of the time integration scheme as figure \ref{fig:Magnitude_velocity_impact_square} proves. However, \acrshort{npc} introduces a minimal energy loss (less than 0.1\%) to significantly reduce the numerical noise. }

\MODIFIED{Although during impact mechanical energy decreases, no real dissipation occurs since after collision the value is again 25 kJ like before collision. Furthermore, kinetic energy has been exactly computed. Hence, this apparently dissipation of mechanical energy is probably due to numerical cancellation of shear terms contribution. }
%%%%%%%%%%%%%%%%
\begin{figure}
%\sidecaption
  \centering
  \resizebox{0.8\hsize}{!}{
\includegraphics[scale=1]{Figures/Energy_square.png} 
  }
  \caption{Snapshots of velocity magnitudes in an elastic square impacted at a point on the right edge. The top row is for \acrshort{npc} updates and the bottom row is for \acrshort{fe} updates.}
  \label{fig:Energy_impact_square}
\end{figure}
%%%%%%%%%%%%%%%%
